Virtualisierung eines Echtzeit-Betriebssystems zur Steuerung eines Roboters mit Schwerpunkt auf die Einhaltung der Echtzeit

### 2024-04-15 Masterarbeit_Robotik_Pamuk.tex
Interrupt Requests Handling

@@

\clearpage
Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency with CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
	\caption[kvm exits]{kvm exits}
	\label{fig:kvm_exit_taskset}
\end{figure}

Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
	\caption[kvm exits default]{kvm exits default}
	\label{fig:kvm_exits_default}
\end{figure}
\clearpage

@@

Figure \ref{fig:gnuplot_max_latency_with_taskset}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_max_latency_with_taskset.png}
	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
	\label{fig:gnuplot_max_latency_with_taskset}
\end{figure}
\clearpage

@@

\chapter{Real-Time Latency Reduction}\label{cha:latency_reduction}
\section{CPU isolation}
Isolating CPUs inolves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process.

Output \ref{output:pse} shows

@@

\clearpage
\section{Kernel tuning}
\clearpage
\section{IRQ affinity}

The IRQ affinity can be configured so that interrupts are handled on CPUs other than CPU 19. This helps reduce the number of kernel threads that are executed on CPU 19.

The script \ref{script:irq_affinity} checks the smp\_affinity for each IRQ on a specific CPU and prints the IRQ numbers that are allowed to run on that CPU, sorted in ascending order.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={Read IRQ affinity per CPU on system},label={script:irq_affinity}]
		#!/bin/bash
		# Get the CPU number from the command-line argument
		CPU=$1
		echo -n "CPU $CPU: "
		for IRQ in /proc/irq/*; do
			if [ -f "$IRQ/smp_affinity" ]; then
				# Read the current smp_affinity
				AFFINITY=$(cat "$IRQ/smp_affinity")
				# Check if the bit for the current CPU is set
				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
					# Print the IRQ number without a newline, followed by a space
					echo -n "${IRQ#/proc/irq/} "
				fi
			fi
		done | sort -n
		# Print a final newline
		echo
\end{lstlisting}
\end{minipage}

@@

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
\begin{lstlisting}[name={Output},label={output:irq_affinity}, breaklines=true]
	sigma_ibo@sigma-ibo:~$ ./check_smp_affinity.sh 19
	CPU 19 IRQ affinity:
	0
	2
	3
	4
	5
	6
	7
	10
	11
	13
	15
	130
	172
	173
	189
	192
\end{lstlisting}
\end{minipage}

@@

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}
\chapter*{\listacroname}
\begin{acronym}[XXXXX]
	\acro{CPU}[CPU]{Central Processing Unit}
	\acro{QEMU}[QEMU]{Quick Emulator}
	\acro{IRQ}[Interrupt Request]{Interrupt Request}
\end{acronym}

>>>>

Figure \ref{fig:max_latency_vapic} shows latency of QEMU vapic Salamander4.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.75\columnwidth]{img/max_latency_vapic.png}
	\caption[Latency taskset]{Latency taskset}
	\label{fig:max_latency_vapic}
\end{figure}

@@

\clearpage
Figure \ref{fig:kvm_exit_taskset} shows \texttt{kvm\_exit} frequency with CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_taskset.png}
	\caption[kvm exits]{kvm exits}
	\label{fig:kvm_exit_taskset}
\end{figure}

Figure \ref{fig:kvm_exits_default} shows \texttt{kvm\_exit} frequency without CPU islation.
\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exits_default.png}
	\caption[kvm exits default]{kvm exits default}
	\label{fig:kvm_exits_default}
\end{figure}

Figure \ref{fig:kvm_exit_vapic} shows \texttt{kvm\_exit} frequency with APIC virtualisation.

\begin{figure}[H]
	\centering
	\includegraphics[width=1.0\columnwidth]{img/kvm_exit_vapic.png}
	\caption[kvm exits default]{kvm exits default}
	\label{fig:kvm_exit_vapic}
\end{figure}
\clearpage

@@

Figure \ref{fig:gnuplot_max_latency_with_taskset}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/with_taskset/gnuplot_max_latency_with_taskset.png}
	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
	\label{fig:gnuplot_max_latency_with_taskset}
\end{figure}

Figure \ref{fig:gnuplot_max_latency_vapic}
\begin{figure}[H]
	\centering
	\includegraphics[width=0.8\columnwidth]{masterthesis-documentation/docs/sigmatek/xenomai/vapic/gnuplot_max_latency_vapic.png}
	\caption[gnuplot latency with taskset]{gnuplot latency with taskset}
	\label{fig:gnuplot_max_latency_vapic}
\end{figure}

@@

\chapter{Real-Time Latency Reduction}\label{cha:latency_reduction}
\section{CPU isolation}
Isolating CPUs involves removing all user-space threads and unbound kernel threads since bound kernel threads are tied to specific CPUs and hence cannot be moved. Also, modifying the \texttt{proc/irq/IRQ\_NUMBER/smp\_affinity} property of each Interrupt IRQ\_NUMBER in the system is part of this process, as described later in section \ref{sec:irq_handling}. Output \ref{output:pse} shows the user and kernel tasks that run on CPU 19. After the isolation, user tasks other than the QEMU process have been removed from running on this CPU. Only few critical kernel threads that are tied to this CPU still take CPU time.

@@

\clearpage
\section{Interrupt Requests Handling}\label{sec:irq_handling}
Once the CPUs were isolated, Interrupt Requests Handling was the next step. Interrupt Requests are used to send a signal to the CPU, prompting it to 'interrupt' its current task and divert its attention to another task. This allows hardware devices to communicate with the CPU through frequent context switches, which can lead to performance degradation, especially in high-performance computing or real-time scenarios. To mitigate this, the IRQs needed to be removed from the isolated CPUs. This was done by manipulating a file in the proc filesystem, namely \texttt{/proc/irq/default\_smp\_affinity}. The value in the \texttt{smp\_affinity} file is a bitmask in hexadecimal format. Each bit in this mask corresponds to a CPU in the system. The least significant bit (LSB) on the right corresponds to the first CPU (CPU0), and the significance increases towards the left until CPU19. In a system with 20 CPUs, if every CPU was reserved for one IRQ, the value for \texttt{smp\_affinity} would be FFFFF. The script in code \ref{script:smp_affinity} was written to check and log the distribution of Interrupt Requests across each CPU in Salamander 4.

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
	\begin{lstlisting}[name={Check distribution of Interrupt Requests across each CPU},label={script:smp_affinity}]
		#!/bin/bash
		# Check if a command-line argument is provided
		if [ -z "$1" ]; then
			echo "Please provide a CPU number as a command-line argument."
			exit 1
		fi
		# Get the CPU number from the command-line argument
		CPU=$1
		# Initialize an empty array to store the IRQ numbers
		IRQs=()
		for IRQ in /proc/irq/*; do
			if [ -f "$IRQ/smp_affinity" ]; then
				# Read the current smp_affinity
				AFFINITY=$(cat "$IRQ/smp_affinity")
				# Check if the bit for the current CPU is set
				if (( (0x$AFFINITY & (1 << CPU)) != 0 )); then
					# Add the IRQ number to the array
					IRQs+=("${IRQ#/proc/irq/}")
				fi
			fi
		done
		# Sort the array
		IFS=$'\n' sorted=($(sort -n <<<"${IRQs[*]}"))
		# Print the CPU number
		echo "CPU $CPU IRQ affinity:"
		# Print the sorted IRQ numbers on separate lines
		for irq in "${sorted[@]}"; do
			echo "$irq"
		done

\end{lstlisting}
\end{minipage}

@@

\vspace{1em}
\begin{minipage}{0.95\columnwidth}
\begin{lstlisting}[name={Output of smp_affinity for CPU 19},label={output:irq_affinity}, breaklines=true]
	sigma_ibo@sigma-ibo:~$ ./check_smp_affinity.sh 19
	CPU 19 IRQ affinity:
	0
	2
	3
	4
	5
	6
	7
	10
	11
	13
	15
	131
	172
	188
	189
	195
\end{lstlisting}
\end{minipage}

By changing the values in

, I could effectively control the assignment of IRQs, ensuring they would not be handled by the isolated CPUs. This manipulation was done carefully, considering the implications and ensuring the stability and integrity of the system.

Through this approach, I was able to create a more streamlined and efficient environment for the execution of specific tasks on the isolated CPUs. By reducing the interruptions caused by IRQs, the isolated CPUs were able to focus more on their assigned tasks, potentially leading to improved performance and more predictable execution times.

\section{Kernel tuning}

@@

\phantomsection
\addcontentsline{toc}{chapter}{\listacroname}
\chapter*{\listacroname}
\begin{acronym}[XXXXX]
	\acro{CPU}[CPU]{Central Processing Unit}
	\acro{QEMU}[QEMU]{Quick Emulator}
	\acro{IRQ}[IRQ]{Interrupt Request}
\end{acronym}
